---
title: "Maternal Effects_Module Overlap"
author: "M.A. Hagadorn"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	warning = TRUE,
	background = "#F7F7F7",
	highlight = TRUE,
	tidy.opts = list(width.cutoff = 60), tidy = TRUE)
```

# Step One: Preparing for Analysis
```{r loadpackages, warning=FALSE, echo=FALSE, message=FALSE, include=FALSE}
#Library Necessary Packages
library(formatR)
library(flashClust)
library(DESeq2) 
library(gplots) 
library(ggplot2)
library(RColorBrewer)
library(affycoretools)
library(genefilter)
library(plotrix)
library(reshape2)
library(vegan)
library(adegenet)
library(stringr)
library(knitr)
library(WGCNA)
library(plyr)
library(dplyr)
library(eulerr)
library(cowplot)
library(ggplot2)
library(gridExtra)
options(stringsAsFactors = FALSE)

```

```{r sessioninformation, echo=FALSE}
installed.packages()[names(sessionInfo()$otherPkgs), "Version"]

```

## Load in RData Files containing necessary files
```{r loadintissueData}

###EGG DATA###
load(file = "../Mrotundata_EGG_ModuleOverlapData.RData")

###OOCTYE DATA###
load(file = "../Mrotundata_OOCYTE_ModuleOverlapData.RData")

```


## Match Labels
```{r matchlabels}
matchedLabels_EGG <- matchLabels(reference = moduleColors_OOCYTE, source = moduleColors_EGG)

#matching labels for egg to the reference(in this case) to oocyte.
#https://www.rdocumentation.org/packages/WGCNA/versions/1.70-3/topics/matchLabels
#The function calculates the overlap of the source and reference modules using Fisher's exact test. It then attempts to relabel source modules such that each source module gets the label of the reference module that it overlaps most with, subject to not renaming two source modules to the same reference module. (If two source modules point to the same reference module, the one with the more significant overlap is chosen.)


oldnew <- data.frame(Oocyte= moduleColors_OOCYTE, old=moduleColors_EGG, new=matchedLabels_EGG)
rownames(oldnew) <- rownames(geneModuleMembership_OOCYTE)
write.table(oldnew, file="ModuleAssignment_OldNewNaming.csv", quote=F, sep=",", 
            row.names=TRUE, col.names=NA)

oldnew_t <- t(oldnew)
```




# Step Two: Module Overlap
## Overlap
```{r moduleoverlap}
#With Matched Lables
overlapTable1 <- overlapTable(labels1 = moduleColors_OOCYTE, labels2 = matchedLabels_EGG)
    #returns a list of components
    #$countTable=a matrix whose rows correspond to modules (unique labels) in lables1 and whose columns correspond to 
                #modules (unique label) in labels 2 --> Gives the number of objects intersecting the two modules

    #$pTable = a matrix whose rows correspond to modules (unique labels) for labels1 and columns correspond to modules 
              #(unique labels) in labels2 --> fishers exact test pvalues for overlat of the two modules


write.table(as.data.frame(overlapTable1$countTable), file="OverlapTableMatched_Count.csv", quote=F, sep=",", 
            row.names=TRUE, col.names=NA)



write.table(as.data.frame(overlapTable1$pTable), file="OverlapTableMatched_P.csv", quote=F, sep=",", 
            row.names=TRUE, col.names=NA)

```


## Assess Overlap for modules of interest
```{r assessmoduleoverlap}
OVERLAP_count <- as.data.frame(overlapTable1$countTable)
OVERLAP_P <- as.data.frame(overlapTable1$pTable)

##EGG modules with significant overlap of oocyte modules of interest
#grey60
ModOvl_grey60 <- names(OVERLAP_P)[which(OVERLAP_P["grey60",] < 0.05)]
grey60_count <- t(OVERLAP_count["grey60", ModOvl_grey60])
grey60_P <- t(OVERLAP_P["grey60", ModOvl_grey60])
grey60 <- cbind(grey60_count, grey60_P)
colnames(grey60) <- c("OL_count", "OL_pValue")

kable(grey60, caption = "Oocyte 'grey60' module overlap with label matched Egg modules")

#midnightblue
ModOvl_midblue <- names(OVERLAP_P)[which(OVERLAP_P["midnightblue",] < 0.05)]
midblue_count <- t(OVERLAP_count["midnightblue", ModOvl_midblue])
midblue_P <- t(OVERLAP_P["midnightblue", ModOvl_midblue])
midblue <- cbind(midblue_count, midblue_P)
colnames(midblue) <- c("OL_count", "OL_pValue")

kable(midblue, caption = "Oocyte 'midnightblue' module overlap with label matched Egg modules")

#violet
ModOvl_violet <-names(OVERLAP_P)[which(OVERLAP_P["violet",] < 0.05)]
violet_count <- t(OVERLAP_count["violet", ModOvl_violet])
violet_P <- t(OVERLAP_P["violet", ModOvl_violet])
violet <- cbind(violet_count, format(round(violet_P, 14), nsmall = 14))
colnames(violet) <- c("OL_count", "OL_pValue")
rownames(violet) <- ModOvl_violet

kable(violet , caption = "Oocyte 'violet' module overlap with label matched Egg modules")


#darkturquoise
ModOvl_darkturquoise <-names(OVERLAP_P)[which(OVERLAP_P["darkturquoise",] < 0.05)]
darkturquoise_count <- t(OVERLAP_count["darkturquoise", ModOvl_darkturquoise])
darkturquoise_P <- t(OVERLAP_P["darkturquoise", ModOvl_darkturquoise])
darkturquoise <- cbind(darkturquoise_count, format(round(darkturquoise_P, 14), nsmall = 14))
colnames(darkturquoise) <- c("OL_count", "OL_pValue")
rownames(darkturquoise) <- ModOvl_darkturquoise

kable(darkturquoise, caption = "Oocyte 'darkturquoise' module overlap with label matched Egg modules")



columnnamesEGG <- c(ModOvl_grey60, "darkgrey", "orange", "blue","darkturquoise","royalblue")

#Empty dataframe
ModofInterest_overlap_OOCYTE <- data.frame(matrix("", ncol=9, nrow=4))
#assign col and row names
rownames(ModofInterest_overlap_OOCYTE) <- c("grey60","midnightblue","violet", "darkturquoise")
colnames(ModofInterest_overlap_OOCYTE) <- columnnamesEGG
#Add specific data points
ModofInterest_overlap_OOCYTE[1,1:4] <- grey60_count
ModofInterest_overlap_OOCYTE[2,c(5,3,4,6)] <- midblue_count
ModofInterest_overlap_OOCYTE[3,3] <- violet_count
ModofInterest_overlap_OOCYTE[4,7:9] <- darkturquoise_count

kable(ModofInterest_overlap_OOCYTE, caption = "Overlap of four modules of interest from the Oocyte data set and the corresponding egg modules (labels matched) that they show significant overlap with")
```

## Pull old labels for correlation comparison
Here, we pull out the old names of the egg modules that have significant overlap with oocyte modules. We are doing this to assess the correlation of these modules with photoperiod
```{r matcholdnameswithnew}
##Egg modules
#Grey60
head(oldnew[which(oldnew$new == "grey60"),2], 1)

#LightSteelBlue1
head(oldnew[which(oldnew$new == "lightsteelblue1"),2], 1)

#MidnightBlue
head(oldnew[which(oldnew$new == "midnightblue"),2], 1)

#cyan
head(oldnew[which(oldnew$new == "cyan"),2], 1)

#DarkGrey
head(oldnew[which(oldnew$new == "darkgrey"),2], 1)

#orange
head(oldnew[which(oldnew$new == "orange"),2], 1)

#darkturquoise
head(oldnew[which(oldnew$new == "darkturquoise"),2], 1)

#Blue
head(oldnew[which(oldnew$new == "blue"),2], 1)

#royalblue
head(oldnew[which(oldnew$new == "royalblue"),2], 1)
```



## Calculating Representation Factor (RF) for overlap
Calculating Representation factors for the Modules with significant overlap.

Representation factor = x/E; E = (n*D)/N

```{r RF}
let <- c("x", "E", " ", "n", "D", "N")
def <- c("# of genes in common between two lists", "Expected # of shared genes", " ", "# of genes in list 1", "# of genes in list 2", "total number of genes")

tab <- cbind(let, def)

library(knitr)
kable(tab, col.names = c("Variable", "Definition"))


#Make a function to calculate RF
RF <- function(n, D, x, GROUP){
  #calculate Expected first
  N <- length(rownames(oldnew))
  E <- (n*D)/N
  
  #calculate overlap
  RF <- x/E
  print(paste("The RF of", GROUP, "=", round(RF, digits=3)))
  return(RF)
}

#OOCYTE--GREY60
#Grey60 vs Grey60
G60.G60 <- RF(length(which(oldnew$Oocyte=="grey60")), 
              length(which(oldnew$new=="grey60")), 
              grey60[rownames(grey60) %in% "grey60",1], 
              "Grey60 & Grey60")

#Grey60 vs LightSteelBlue1
G60.LSB1 <- RF(length(which(oldnew$Oocyte=="grey60")), 
               length(which(oldnew$new=="lightsteelblue1")), 
               grey60[rownames(grey60) %in% "lightsteelblue1",1], 
               "Grey60 & LightSteelBlue1")

#Grey60 vs MidnightBlue
G60.MB <- RF(length(which(oldnew$Oocyte=="grey60")), 
               length(which(oldnew$new=="midnightblue")),
               grey60[rownames(grey60) %in% "midnightblue",1], 
             "Grey60 & MidnightBlue")

#Grey60 vs Cyan
G60.cyan <- RF(length(which(oldnew$Oocyte=="grey60")), 
               length(which(oldnew$new=="cyan")),
               grey60[rownames(grey60) %in% "cyan",1], 
               "Grey60 & Cyan")



#OOCYTE--MidnightBlue
#MidnightBlue & DarkGrey
MB.DG <- RF(length(which(oldnew$Oocyte=="midnightblue")), 
               length(which(oldnew$new=="darkgrey")),
               midblue[rownames(midblue) %in% "darkgrey",1], 
            "MidnightBlue & DarkGrey")

#MidnightBlue & LightSteelBlue1
MB.LSB1 <- RF(length(which(oldnew$Oocyte=="midnightblue")), 
               length(which(oldnew$new=="lightsteelblue1")),
               midblue[rownames(midblue) %in% "lightsteelblue1",1], 
              "MidnightBlue & LightSteelBlue1")

#MidnightBlue & MidnightBlue
MB.LSB1 <- RF(length(which(oldnew$Oocyte=="midnightblue")), 
               length(which(oldnew$new=="midnightblue")),
               midblue[rownames(midblue) %in% "midnightblue",1], 
              "MidnightBlue & MidnightBlue")

#MidnightBlue & Orange
MB.OG <- RF(length(which(oldnew$Oocyte=="midnightblue")), 
               length(which(oldnew$new=="orange")),
               midblue[rownames(midblue) %in% "orange",1], 
            "MidnightBlue & Orange")



#OOCYTE--Violet
#MidnightBlue & Violet
VI.LSB1 <- RF(length(which(oldnew$Oocyte=="violet")), 
               length(which(oldnew$new=="lightsteelblue1")),
               as.numeric(noquote(violet[rownames( violet) %in% "lightsteelblue1",1])), 
              "Violet & LightSteelBlue1")


#OOCYTE--DarlTurquoise
#DarkTurquoise vs DarkTurquoise
DT.DT <- RF(length(which(oldnew$Oocyte=="darkturquoise")), 
               length(which(oldnew$new=="darkturquoise")),
               as.numeric(noquote(darkturquoise[rownames(darkturquoise) %in% "darkturquoise",1])),
                          "DarkTurquoise & DarkTurquoise")


#DarkTurquoise vs DarkTurquoise
DT.BL <- RF(length(which(oldnew$Oocyte=="darkturquoise")), 
               length(which(oldnew$new=="blue")),
               as.numeric(noquote(darkturquoise[rownames(darkturquoise) %in% "blue",1])),
                          "DarkTurquoise & Blue")

DT.RB <- RF(length(which(oldnew$Oocyte=="darkturquoise")), 
               length(which(oldnew$new=="royalblue")),
               as.numeric(noquote(darkturquoise[rownames(darkturquoise) %in% "royalblue",1])),
                          "DarkTurquoise & RoyalBlue")

```


## Proportion overlap of oocyte module and egg module
```{r overlap}

#Grey60 vs #Grey60
#Overlap prop with oocyte
grey60["grey60",1]/length(which(oldnew$Oocyte=="grey60"))
#Overlap prop with egg
grey60["grey60",1]/length(which(oldnew$new=="grey60"))

#Grey60 vs LightSteelBlue1
#Overlap prop with oocyte
grey60["lightsteelblue1",1]/length(which(oldnew$Oocyte=="grey60"))
#Overlap prop with egg
grey60["lightsteelblue1",1]/length(which(oldnew$new=="lightsteelblue1"))


#Grey60 vs MidnightBlue
#Overlap prop with oocyte
grey60["midnightblue",1]/length(which(oldnew$Oocyte=="grey60"))
#Overlap prop with egg
grey60["midnightblue",1]/length(which(oldnew$new=="midnightblue"))

#Grey60 vs Cyan
#Overlap prop with oocyte
grey60["cyan",1]/length(which(oldnew$Oocyte=="grey60"))
#Overlap prop with egg
grey60["cyan",1]/length(which(oldnew$new=="cyan"))


#OOCYTE--MidnightBlue
#MidnightBlue & DarkGrey
#Overlap prop with oocyte
midblue["darkgrey",1]/length(which(oldnew$Oocyte=="midnightblue"))
#Overlap prop with egg
midblue["darkgrey",1]/length(which(oldnew$new=="darkgrey"))

#MidnightBlue & LightSteelBlue1
#Overlap prop with oocyte
midblue["lightsteelblue1",1]/length(which(oldnew$Oocyte=="midnightblue"))
#Overlap prop with egg
midblue["lightsteelblue1",1]/length(which(oldnew$new=="lightsteelblue1"))

#MidnightBlue & MidnightBlue
#Overlap prop with oocyte
midblue["midnightblue",1]/length(which(oldnew$Oocyte=="midnightblue"))
#Overlap prop with egg
midblue["midnightblue",1]/length(which(oldnew$new=="midnightblue"))

#MidnightBlue & Orange
#Overlap prop with oocyte
midblue["orange",1]/length(which(oldnew$Oocyte=="midnightblue"))
#Overlap prop with egg
midblue["orange",1]/length(which(oldnew$new=="orange"))



#OOCYTE--Violet
#Violet & LightSteelBlue1
#Overlap prop with oocyte
as.numeric(violet["lightsteelblue1",1])/length(which(oldnew$Oocyte=="violet"))
#Overlap prop with egg
as.numeric(violet["lightsteelblue1",1])/length(which(oldnew$new=="lightsteelblue1"))


#OOCYTE--DarkTurquoise
#DarkTurquoise vs DarkTurquoise
#Overlap prop with oocyte
as.numeric(darkturquoise["darkturquoise",1])/length(which(oldnew$Oocyte=="darkturquoise"))
#Overlap prop with egg
as.numeric(darkturquoise["darkturquoise",1])/length(which(oldnew$new=="darkturquoise"))


#DarkTurquoise vs Blue
#Overlap prop with oocyte
as.numeric(darkturquoise["blue",1])/length(which(oldnew$Oocyte=="darkturquoise"))
#Overlap prop with egg
as.numeric(darkturquoise["blue",1])/length(which(oldnew$new=="blue"))

#DarkTurquoise vs Roya1Blue
#Overlap prop with oocyte
as.numeric(darkturquoise["royalblue",1])/length(which(oldnew$Oocyte=="darkturquoise"))
#Overlap prop with egg
as.numeric(darkturquoise["royalblue",1])/length(which(oldnew$new=="royalblue"))
```





## Proportion overlap of oocyte module with all genes of significantly overlapped Egg modules
```{r overlap_ALL}

#Grey60
sum(grey60[,1])/length(which(oldnew$Oocyte=="grey60"))

#MidnightBlue
sum(midblue[,1])/length(which(oldnew$Oocyte=="midnightblue"))

#Violet
sum(as.numeric(violet[,1]))/length(which(oldnew$Oocyte=="violet"))

#DarkTurquoise
sum(as.numeric(darkturquoise[,1]))/length(which(oldnew$Oocyte=="darkturquoise"))

```
```{r save}
ModofInterest_overlap_OOCYTE_numericoocytemoduleid <- ModofInterest_overlap_OOCYTE
rownames(ModofInterest_overlap_OOCYTE_numericoocytemoduleid) <- c("oocyteMod16", "oocyteMod4", "oocyteMod29", "oocyteMod20")
colnames(ModofInterest_overlap_OOCYTE_numericoocytemoduleid) <- c("eggMod20", "eggMod25", "eggMod4", "eggMod17", "eggMod13", "eggMod15", "eggMod5", "eggMod24", "eggMod23")

write.table(ModofInterest_overlap_OOCYTE_numericoocytemoduleid, "G:/DesktopDocuments/Documents/PHD/Manuscripts/Bee/Mrot_MaternalEffects/WGCNA/MRot_WGCNA/ModuleOverlap/OverlapTables/ModofInterest_overlapwithOOCYTE_numericmoduleid.csv", sep=",", col.names = TRUE, row.names=TRUE)

write.table(ModofInterest_overlap_OOCYTE, "G:/DesktopDocuments/Documents/PHD/Manuscripts/Bee/Mrot_MaternalEffects/WGCNA/MRot_WGCNA/ModuleOverlap/OverlapTables/ModofInterest_overlapwithOOCYTE_originalnamingbycolors.csv", sep=",", col.names = TRUE, row.names=TRUE)

save(ModofInterest_overlap_OOCYTE, oldnew, overlapTable1, ModofInterest_overlap_OOCYTE_numericoocytemoduleid, ModofInterest_overlap_OOCYTE_numericoocytemoduleid, file = "Mrotundata_BuildEulerPlotInformation")

```