---
title: 'WGCNA: Megachile rotundata Maternal Effects Project: Oocyte Samples'
author: "M.A. Hagadorn"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	warning = TRUE,
	background = "#F7F7F7",
	highlight = TRUE,
	tidy.opts = list(width.cutoff = 60), tidy = TRUE)
```

# Step One: Preparing for Analysis
```{r loadpackages, warning=FALSE, echo=FALSE, message=FALSE, include=FALSE}
#Library Necessary Packages
library(formatR)
library(flashClust)
library(DESeq2) 
library(gplots) 
library(ggplot2)
library(RColorBrewer)
library(affycoretools)
library(genefilter)
library(plotrix)
library(reshape2)
library(vegan)
library(adegenet)
library(stringr)
library(WGCNA)
library(plyr)
options(stringsAsFactors = FALSE)

```

```{r sessioninformation, echo=FALSE}
installed.packages()[names(sessionInfo()$otherPkgs), "Version"]

```

Prior to the start of this analysis, 852 genes were removed from the analysis. This is based on WGCNA gene removal criteria applied to *ALL* samples. As such, the data frames containing **egg** and **ooctye** seperately will have equal number genes for downstream analyses. This is important for module overlap assessment.

So, first we load in the oocyte specific expression data (that has been cleaned, i.e. genes removed).
```{r loadOOCYTEData, echo=FALSE}
load("data/Mrotundata_expdataClean_OOCYTE.RData")
```

# Step One: cluster samples
## Cluster samples  to determine if there are obvious outliers
```{r clustersamples}
#create a distance matrix and do a hierachical cluster analysis
#using data set with genes removed
sampleTree_precut <- hclust(dist(expd_OOCYTE), method = "average")

#plot the results
##set parameters
par(cex=0.6, mar=c(0,4,2,0))

##plot
plot(sampleTree_precut,
     main="OOCYTE SAMPLES: clustering to detect outliers", 
     sub="", xlab="", cex.lab =1.5, cex.axis=1.5, cex.main=2, cex=0.9)

###
###
###
#NO OUTLIERS DETECTED

expdatOOCYTE <- expd_OOCYTE
```


## Read in the trait data the correspond to the samples
```{r loadtraitdata}
sampleData_all <- read.csv2("data/sample_data.csv", header = TRUE, sep= ",", as.is = TRUE)
dim(sampleData_all)
names(sampleData_all)

##Remove columns that hold information we do not need to analysis
sampleData <- sampleData_all[sampleData_all$oocyte_egg=="oocyte",-c(4,7,8)]
dim(sampleData)
names(sampleData)


#Make new dataframe analogous to expression data that will hold the traits
samples <- rownames(expdatOOCYTE) #save names of all rows in variable expdatOOCYTE_cleaned
variableRows <- match(samples,sampleData$id) #returns vector of positions for these

varData <- sampleData[variableRows, -1] #make new dataframe with ID removed but positions as rows
rownames(varData) <- sampleData[variableRows, 1] #assign sample name as row name instead of position

collectGarbage() #memory thing
```

## Vizualization: sample tree and traits
```{r vizualize, fig.height = 16, fig.width = 14}
#Setup:

##expression data w/ genes and outliers removed==stored as expdatOOCYTE
##Sample variables==stored as VarData


#re-cluster samples (because we removed the one sample)
sampleTree <- hclust(dist(expdatOOCYTE), method="average")


#convert traits to a color representation: white=low; red=high, grey=missing
varColors <- labels2colors(varData, naColor = "grey")


#plot
plotDendroAndColors(sampleTree, varColors,  #Sample tree with all samples
                    groupLabels = names(varData), 
                    setLayout=TRUE, 
                    autoColorHeight = TRUE, 
                    colorHeightMax = 0.3,
                    main="Sample dendrogram and trait heatmap")
```


```{r save, echo=FALSE}
## Save expression and trait data
save(expdatOOCYTE, varData, file="Mrotundata_OOCYTE_step1_expression&variablesetup.RData")
```




# Step Two: Gene network and module identification
## Soft-threshold
```{r softy}
#chose a set of soft-thresholding powers
#these are candidate powers to use for discovering the power that fits the best
powers <- c(c(1:10), seq(from=12, to=30, by=2))

#call network topology analysis function
softThres <- pickSoftThreshold(expdatOOCYTE, powerVector = powers, verbose = 5)

#Plot the results to determine the best index
par(mfrow= c(1,2), mar=c(0,4,2,0))
#scale-free topology fit index based on the soft-thresholding power

##notes on sign() function used below
  #returns vector w/ signs of corresponding elements of x (the sign of a real             
  #number is 1, 0, or -1 if the number is positive, zero, or negative, respectively)
plot(softThres$fitIndices[,"Power"], -sign(softThres$fitIndices[,"slope"])*softThres$fitIndices[,"SFT.R.sq"],
     xlab="Soft Threshold (power)",
     ylab="Scale Free Topology Model Fit, signed R^2",
     type="n", #"n" for no plotting
     main="Scale Independence",
     ylim=c(-0.6, 1.0))
#add text
text(softThres$fitIndices[,"Power"], -sign(softThres$fitIndices[,"slope"])*softThres$fitIndices[,"SFT.R.sq"],
     labels=powers, cex=0.9, col="red")

#h=R^2 cutoff
abline(h=0.9, col="blue")
abline(h=0.8, col="green")

#Mean connectivity as a function of the soft-threshold power
plot(softThres$fitIndices[,"Power"], softThres$fitIndices[,"mean.k."],
     xlab="Soft Threshold (power)",
     ylab="Mean Connectivity",
     type="n",
     main="Mean Connectivity")
text(softThres$fitIndices[,"Power"], softThres$fitIndices[,"mean.k."], labels=powers, cex=0.9, col="red")    



#See: https://horvath.genetics.ucla.edu/html/CoexpressionNetwork/Rpackages/WGCNA/faq.html for details about connectivity

#Selecting a power of 4; BUT we type (i.e. oocyte vs egg are likely strong drivers for samples)
```


## Co-expression similarity and adjacency
```{r sim and adjacency}
softPower <- 4 #decided via the step above
adjacency <- adjacency(expdatOOCYTE, power=softPower) 
```


## TOM-Topological Overlap Matrix
```{r TOM}
#to minimize effects of noise and spurious associations, transform adjacency into TOM
TOM <- TOMsimilarity(adjacency) 
  #Calculates Topological Overlap Matrix

dissTOM <- 1-TOM
  #calculates dissimilarity
```



## Clustering using TOM
```{r hclustTOM, fig.height = 10, fig.width = 14}
# Call the hierarchical clustering function
geneTree <- hclust(as.dist(dissTOM), method = "average")

# Plot the resulting clustering tree (dendrogram)
plot(geneTree, 
     xlab="", 
     sub="", 
     main = "Gene clustering on TOM-based dissimilarity",
     labels = FALSE, 
     hang = 0.04)
```


## Assigning modules and vizualize
```{r modules, fig.height = 16, fig.width = 14}
# Identification of the modules
dynamicModules <- cutreeDynamic(dendro = geneTree, 
                             distM = dissTOM,
                             deepSplit = 2, 
                             pamRespectsDendro = FALSE,
                             minClusterSize = 30) #Large modules=high minimum
#information suggests this minimum is good?

table(dynamicModules) #view number of modules and genes count assigned to each
#49 modules

#convert numeric module labels to into colors
dynamicColors <- labels2colors(dynamicModules)

#plot
plotDendroAndColors(geneTree,
                    dynamicColors,
                    groupLabels = "Dynamic Tree Cuts",
                    dendroLabels = FALSE,
                    hang=0.03,
                    addGuide=TRUE,
                    guideHang = 0.05,
                    main="Gene dendrogram and module colors")

```



## Merge modules with very similar expression profiles
```{r mergemodules, fig.height = 16, fig.width = 14}

# Calculate eigengenes values
ModuleEigenLs <- moduleEigengenes(expdatOOCYTE, colors = dynamicColors)

#eigengenes of modules in DF form; colum corresponds to one eigengene
#eigengene=1st principal component of a module expression matrix
MEs <- ModuleEigenLs$eigengenes 


# Calculate dissimilarity of module eigengenes
MEDiss <- 1-cor(MEs, use='pairwise.complete.obs')
MEDiss_noNA <- subset(MEDiss, select=-c(MEgrey)) #MEgrey has NaN--divided by zero
MEDiss_noNA <- MEDiss_noNA[!(row.names(MEDiss_noNA) %in% "MEgrey"),]

# Cluster module eigengenes
  #create dist matrix based on MEDiss and then h clust
METree <- hclust(as.dist(MEDiss_noNA), method = "average") 

# Plot the result
plot(METree,
     main = "Clustering of module eigengenes",
     xlab = "",
     sub = "")



##Merge Selection Threshold
#This corresponds to a similarity of .70 to merge a module
MEDissThres <- 0.25

# Plot the cut line into the dendrogram
abline(h=MEDissThres, col = "blue")

# Call an automatic merging function
mergeModules <- mergeCloseModules(expdatOOCYTE, dynamicColors, cutHeight = MEDissThres, verbose = 3) #cutoff of .25

# The merged module colors
mergedColors <- mergeModules$colors  #colors for .25

# Eigengenes of the new merged modules:
mergedMEs <- mergeModules$newMEs
```


## Vizualization: old modules vs new, merged modules
```{r viz-newold, fig.height = 16, fig.width = 14}
#plot new and old
plotDendroAndColors(geneTree, 
                    cbind(dynamicColors, mergedColors),
                    c("Dynamic Tree Cut", "Merged dynamic 0.25"),
                    dendroLabels = FALSE, 
                    hang = 0.03,
                    addGuide = TRUE, 
                    guideHang = 0.05)

```


## Save relevant information
```{r savemodules}
# Rename to moduleColors
moduleColors <- mergedColors

# Construct numerical labels corresponding to the colors
colorOrder <- c("grey", standardColors(50))
moduleLabels <- match(moduleColors, colorOrder)-11

MEs <- mergedMEs

# Save module colors and labels for use in subsequent parts
save(MEs, moduleLabels, moduleColors, geneTree, TOM, dissTOM, 
     file = "Mrotundata_OOCYTE_step2_ModuleAssignment&Merge.RData") 
```


## Quantify module-variable associations
```{r calculateP}
# Define numbers of genes and samples
nGenes <- ncol(expdatOOCYTE)
nSamples <- nrow(expdatOOCYTE)

#Recalculate Module Eigen with color labels
MEs_modVar <- moduleEigengenes(expdatOOCYTE, moduleColors)$eigengenes

#Reorder the eigen vectors such that more correlated ones are next to each other
MEs <- orderMEs(MEs_modVar)

#Identify significant variable associations
#correlation=linear relationship between two variables

#######Turning variables into binary values
varData_bi <- varData[,-c(1,5)]

#TX
varData_bi$tx <- gsub("17U1", 1, varData_bi$tx)
varData_bi$tx <- gsub("17U14", 14, varData_bi$tx)

#cage
varData_bi$cage <- gsub(9, 1, varData_bi$cage)
varData_bi$cage <- gsub(10, 2, varData_bi$cage)


moduleVarCor <- cor(MEs, varData_bi, use="p")

#calculate pvalue given the correlation
moduleVarPValue <- corPvalueStudent(moduleVarCor, nSamples)

#Adjustment for pvalue
nmodules <- ncol(MEs)
nvars <- ncol(varData_bi)

adj_P <- 0.05/(nmodules+nvars)
adj_P
```


## Vizualizae module-variable associations
```{r Vizualize_m-t, fig.height=12, fig.width=8}
#Display correlations and p-values
textMatrix <- paste(signif(moduleVarCor, 2),
                    "\n(",
                    signif(moduleVarPValue, 1),
                    ")",
                    sep = "")

textMatDim <- dim(moduleVarCor)

##pdf("hm_25.pdf", height = 12, width=8)

par(mar=c(5,9,2,1))
#Use heatmap to display correlation values in the heat map
#table is color-coded according to correlation 
labeledHeatmap(Matrix=moduleVarCor,
               xLabels = names(varData_bi),
               yLabels = names(MEs),
               ySymbols = names(MEs),
               colorLabels = FALSE,
               colors=blueWhiteRed(40),
               textMatrix = textMatrix,
               setStdMargins = FALSE,
               cex.text = 0.7,
               zlim = c(-1,1),
               main = "OOCYTE TISSUE: Module-variable relationships cutHeight = 0.25")

#dev.off()
```
For significance, adjusted p-value threshold should be `adj_P`.

## Experimental Factors and Significant Modules
For each gene, we define "fuzzy" measures of module membership by correlating gene expression with module eigengene of a given module.

If Module Membership (MM) for a particular module is close to 0, the gene is not part of that module.
If MM of a particular module is  close to 1 or -1, it is highly connected to that particular module.
Sign = positive or negative relationship

### Treatment Group
```{r tx_grouping}
#assigning the MM can be done for all genes, irrespective of what their original module membership is

# Define variable weight containing the weight column of varData_bi
# tx per samples
tx <- as.data.frame(varData_bi$tx)

#Add column header
names(tx) <- "tx"

# names (colors) of the modules
moduleNames <- substring(names(MEs_modVar), 3)

#identifying gene module membership based on gene and Module Eigengene
#correlation
#0=not in that module
#closer to 1 and -1 = highly connected to that module
geneModuleMembership <- as.data.frame(cor(expdatOOCYTE, MEs_modVar, use = "p"))

#Module Membership p-value
#pvalue for likelihood of being in a particular module
MMPvalue <- as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples))

#adding names
names(geneModuleMembership) <- paste("MM", moduleNames, sep="") #adding MM to module name
names(MMPvalue) <- paste("p.MM", moduleNames, sep="")


geneVarSignificance_tx <- as.data.frame(cor(expdatOOCYTE, tx, use = "p"))

#Identifying gene significance
#higher = more biologically relevant
GSPvalue_tx <- as.data.frame(corPvalueStudent(as.matrix(geneVarSignificance_tx), nSamples))

#Adding names
names(geneVarSignificance_tx) <- paste("GS.", names(tx), sep="")
names(GSPvalue_tx) <- paste("p.GS.", names(tx), sep="")

print(subset(GSPvalue_tx, GSPvalue_tx$p.GS.tx < 0.01))

print(subset(GSPvalue_tx, GSPvalue_tx$p.GS.tx < 0.001))
```


### Wave Group
```{r wave}
#assigning the MM can be done for all genes, irrespective of what their original module membership is

# Define variable weight containing the weight column of varData_bi
# tx per samples
wave <- as.data.frame(varData_bi$wave)

#Add column header
names(wave) <- "wave"

# names (colors) of the modules
#moduleNames <- substring(names(MEs_modVar), 3)

#identifying gene module membership based on gene and Module Eigengene
#correlation
#0=not in that module
#closer to 1 and -1 = highly connected to that module
#geneModuleMembership <- as.data.frame(cor(expdatOOCYTE, MEs_modVar, use = "p"))

#Module Membership p-value
#pvalue for likelihood of being in a particular module
#MMPvalue <- as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples))

#adding names
#names(geneModuleMembership) <- paste("MM", moduleNames, sep="") #adding MM to module name
#names(MMPvalue) <- paste("p.MM", moduleNames, sep="")


geneVarSignificance_wave <- as.data.frame(cor(expdatOOCYTE, wave, use = "p"))

#Identifying gene significance
#higher = more biologically relevant
GSPvalue_wave <- as.data.frame(corPvalueStudent(as.matrix(geneVarSignificance_wave), nSamples))

#Adding names
names(geneVarSignificance_wave) <- paste("GS.", names(wave), sep="")
names(GSPvalue_wave) <- paste("p.GS.", names(wave), sep="")

print(subset(GSPvalue_wave, GSPvalue_wave$p.GS.wave < 0.01))

print(subset(GSPvalue_wave, GSPvalue_wave$p.GS.wave < 0.001))
```

#### Intramodular Analysis
##### Grey60 Module
```{r ModuleGSandMM, fig.height = 7, fig.width = 7, fig.align="center"}
#asign module
moduleofinterest1 <- "grey60"

#saying get "brown" location from vector modNames
grey60 <- match(moduleofinterest1, moduleNames)

#identifying which genes fall into this module as T or F
moduleGenes_grey60 <- moduleColors==moduleofinterest1


par(mfrow = c(1,1))
#scatter plot annotated with correlation, pvalue, and regression lien
verboseScatterplot(abs(geneModuleMembership[moduleGenes_grey60, grey60]), #selects dark green only
                   abs(geneVarSignificance_wave[moduleGenes_grey60, 1]), #selects oocyte_egg only
                   xlab = paste("Module Membership in", moduleofinterest1, "module"),
                   ylab = "Gene significance for Wave",
                   main = paste("Module membership vs. gene significance\n"),
                   cex.main = 1.2, 
                   cex.lab = 1.2, 
                   cex.axis = 1.2, 
                   col = moduleofinterest1)
```


##### Midnight Blue
```{r ModuleGSandMM_2, fig.height = 7, fig.width = 7, fig.align="center"}
#asign module
moduleofinterest2 <- "midnightblue"

#saying get "brown" location from vector modNames
midnightblue <- match(moduleofinterest2, moduleNames)

#identifying which genes fall into this module as T or F
moduleGenes_midnightblue <- moduleColors==moduleofinterest2


par(mfrow = c(1,1))
#scatter plot annotated with correlation, pvalue, and regression lien
verboseScatterplot(abs(geneModuleMembership[moduleGenes_midnightblue, midnightblue]), #selects midnightblue only
                   abs(geneVarSignificance_wave[moduleGenes_midnightblue, 1]), #selects oocyte_egg only
                   xlab = paste("Module Membership in", moduleofinterest2, "module"),
                   ylab = "Gene significance for Wave",
                   main = paste("Module membership vs. gene significance\n"),
                   cex.main = 1.2, 
                   cex.lab = 1.2, 
                   cex.axis = 1.2, 
                   col = moduleofinterest2)
```

##### Violet Module
```{r ModuleGSandMM_3, fig.height = 7, fig.width = 7, fig.align="center"}
#asign module
moduleofinterest3 <- "violet"

#saying get "brown" location from vector modNames
violet <- match(moduleofinterest3, moduleNames)

#identifying which genes fall into this module as T or F
moduleGenes_violet <- moduleColors==moduleofinterest3


par(mfrow = c(1,1))
#scatter plot annotated with correlation, pvalue, and regression lien
verboseScatterplot(abs(geneModuleMembership[moduleGenes_violet, violet]), #selects black only
                   abs(geneVarSignificance_wave[moduleGenes_violet, 1]), #selects oocyte_egg only
                   xlab = paste("Module Membership in", moduleofinterest3, "module"),
                   ylab = "Gene significance for Wave",
                   main = paste("Module membership vs. gene significance\n"),
                   cex.main = 1.2, 
                   cex.lab = 1.2, 
                   cex.axis = 1.2, 
                   col = moduleofinterest3)
```



## Creat Gene and Module Information Dataframe
### Make DF
```{r geneinfodat}
# Create the starting data frame
geneInfo_raw <- data.frame(geneid = names(expdatOOCYTE),
                              moduleColor = moduleColors,
                              geneVarSignificance_tx=geneVarSignificance_tx,
                              GSPvalue_tx= GSPvalue_tx,
                              geneVarSignificance_wave=geneVarSignificance_wave,
                              GSPvalue_wave= GSPvalue_wave)

# Order modules by their significance for treatment group
moduleOrder <- order(-abs(cor(MEs, tx, use = "p")))

# Add module membership information in the chosen order
for (module in 1:ncol(geneModuleMembership))
{
  oldNames <- names(geneInfo_raw)
  geneInfo_raw <- data.frame(geneInfo_raw, geneModuleMembership[, moduleOrder[module]], MMPvalue[, moduleOrder[module]]);
  names(geneInfo_raw) <- c(oldNames, paste("MM.", moduleNames[moduleOrder[module]], sep=""), paste("p.MM.", moduleNames[moduleOrder[module]], sep=""))
}


# Order the genes in the geneInfo variable first by module color, then by geneVarSignificance
geneOrder <- order(geneInfo_raw$moduleColor, -abs(geneInfo_raw$GS.tx));
geneInfo_OOCYTE = geneInfo_raw[geneOrder, ]

```

### Module Frequencies
```{r module counts graph, fig.height=10, fig.width=8}

modulecounts <- count(geneInfo_OOCYTE, 'moduleColor')

modulecounts_ordered <- modulecounts[order(-modulecounts$freq),]

par(mgp=c(1,1,0), mar=c(2.5,8.5,0,1))
barplot2(modulecounts_ordered$freq, horiz = TRUE, names.arg = 
           modulecounts_ordered$moduleColor, cex.axis = 1, cex.names = 1, las=1, 
           col=modulecounts_ordered$moduleColor, xlim = c(0,5000), xaxt="n")
axis(1, pos=0, at = seq(0,5000,500)) # Draw x-axis at 0 line
mtext(text = "Module Assignment", side = 2, line = 7, cex = 1.4)
mtext(text = "Gene Frequency", side = 1, line = 1, cex = 1.4)

```



## Write Gene Information to csv
```{r geneinfo_CSV}
# Save module colors and labels for use in subsequent parts
save(geneModuleMembership, geneVarSignificance_tx, geneVarSignificance_wave, GSPvalue_tx, GSPvalue_wave, wave, tx, file = "Mrotundata_OOCYTE_step3_ModuleMembership&GeneInfoID.RData")



write.csv(geneInfo_OOCYTE, file = "geneInfo_OOCYTE.csv")
```


## Saving RData for Module Overlap
```{r rdata_moduleoverlap}
###RENAME FILES CONTAINING THE OOCYTE LABEL###
#Makes life easier for module overlap loadin
geneModuleMembership_OOCYTE <- geneModuleMembership
MEs_OOCYTE <- MEs
moduleColors_OOCYTE <- moduleColors
moduleLabels_OOCYTE <- moduleLabels


# Save module colors and labels for use in subsequent parts
save(geneModuleMembership_OOCYTE,moduleColors_OOCYTE, moduleLabels_OOCYTE, MEs_OOCYTE, file = "Mrotundata_OOCYTE_ModuleOverlapData.RData")
```
